---
title: "Setting a default rounding specification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting a default rounding specification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

If you ever get tired of having to write `peri_glue(..., rspec = your_rspec)` and would rather just write `peri_glue(...)`, you should find this vignette helpful.

```{r setup}
library(periglue)
```

## Setting a default with `set_default_rounding_spec()`

Build a rounding specification with `round_spec()` and the `round_`/`format_` helper families, then pass it to `set_default_rounding_spec()`. All subsequent calls to `peri_glue()` and `peri_value()` will use it automatically.

```{r}
rspec <- round_spec() |>
  round_using_decimal(digits = 5)

set_default_rounding_spec(rspec)

peri_value(pi)
```

## Clearing the default

Pass `NULL` to revert to factory settings.

```{r}
set_default_rounding_spec(NULL)

peri_value(pi)
```

## Saving and restoring

`set_default_rounding_spec()` invisibly returns the *previous* default, which makes it straightforward to save and restore around a block of code.

```{r}
rspec <- round_spec() |> round_using_decimal(digits = 3)

old <- set_default_rounding_spec(rspec)
peri_value(pi)

# Restore whatever was set before
set_default_rounding_spec(old)
peri_value(pi)
```

## Factory defaults vs. stored defaults

`round_spec(force_default = TRUE)` bypasses both the stored default and any `periglue.*` global options, always returning the factory settings. This is useful if you need a clean baseline to build from regardless of session state.

```{r}
set_default_rounding_spec(round_spec() |> round_using_decimal(digits = 5))

# Ignores the stored default
round_spec(force_default = TRUE)$digits
```

```{r, include = FALSE}
set_default_rounding_spec(NULL)
```
