---
title: "Migrating from table.glue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Migrating from table.glue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`periglue` is a fork of `table.glue` with new features, bug fixes, and a
cleaner API. This vignette covers what changed and how to update code that
used `table.glue`.

```{r setup}
library(periglue)
```

## Renamed functions

All `table_*` functions are now `peri_*`:

| `table.glue`    | `periglue`      |
|:----------------|:----------------|
| `table_glue()`  | `peri_glue()`   |
| `table_value()` | `peri_value()`  |
| `table_pvalue()`| `peri_pvalue()` |
| `table_ester()` | `peri_ester()`  |
| `table_estin()` | `peri_estin()`  |

All other functions (`round_spec()`, `round_using_*()`, `format_*()`,
`bracket_*()`, `as_inline()`) are unchanged.

## New default rounding

`table.glue` used magnitude-based rounding by default — the number of decimal
places depended on the size of the value:

- values < 1 → 2 decimal places
- values 1–10 → 1 decimal place
- values > 10 → 0 decimal places

This meant a vector of values could be formatted inconsistently:

```{r}
# Simulated mean ± SD pairs at different scales
peri_value(c(0.82, 4.7, 31.2), rspec = round_spec() |> round_using_magnitude())
```

`periglue` defaults to **1 decimal place for all values**, which is more
predictable and appropriate for most table cells:

```{r}
peri_value(c(0.82, 4.7, 31.2))
```

If you preferred the old magnitude-based behaviour, pass an explicit `rspec`:

```{r}
rspec_mag <- round_spec() |>
  round_using_magnitude(digits = c(2, 1, 0), breaks = c(1, 10, Inf))

peri_value(c(0.82, 4.7, 31.2), rspec = rspec_mag)
```

## New way to set a session default

`table.glue` required users to rename rspec fields manually before passing
them to `options()`:

```r
# table.glue approach — no longer recommended
rspec <- round_spec() |> round_using_decimal(digits = 2)
names(rspec) <- paste("periglue", names(rspec), sep = ".")
options(rspec)
```

This worked but was easy to get wrong, and CRAN discourages packages from
modifying user options.

`periglue` provides `set_default_rounding_spec()`, which stores the
specification in the package's own internal environment rather than in global
options:

```{r}
rspec <- round_spec() |> round_using_decimal(digits = 2)
set_default_rounding_spec(rspec)

peri_value(pi)
peri_glue("mean (SD): {mean(mtcars$mpg)} ({sd(mtcars$mpg)})")
```

Pass `NULL` to clear the default and revert to factory settings:

```{r}
set_default_rounding_spec(NULL)
peri_value(pi)
```

`set_default_rounding_spec()` returns the previous default invisibly, making it
easy to save and restore around a block of code:

```{r}
rspec_narrow <- round_spec() |> round_using_decimal(digits = 3)

old <- set_default_rounding_spec(rspec_narrow)
peri_value(pi)

set_default_rounding_spec(old)  # restore whatever was set before
peri_value(pi)
```

## New function: `peri_n_pct()`

The most common cell format in demographic tables — a count with its
percentage — previously required `peri_glue()`:

```{r}
n   <- c(48, 32, 20)
pct <- c(0.48, 0.32, 0.20)

# old approach — relies on n and pct being in scope
peri_glue("{n} ({round(100 * pct, 1)}%)")
```

This works but rounds the percentage using the same specification as everything
else in the template, and gives no control over formatting the count separately
from the percentage. `peri_n_pct()` handles both in one call:

```{r}
peri_n_pct(n = n, pct = pct)
```

Counts are always displayed as whole numbers regardless of the `rspec` used for
the percentage, and the formatting settings (big mark, decimal mark,
`miss_replace`) are inherited from a single specification:

```{r}
peri_n_pct(n = 12480, pct = 0.621)
```

Missing percentages — common when a subgroup denominator is zero — are handled
cleanly:

```{r}
peri_n_pct(n = c(10, 0), pct = c(0.10, NaN))
```

The output format is customisable via the `form` argument:

```{r}
peri_n_pct(n = 48, pct = 0.48, form = "{n}/{100} [{pct}%]")
```

## Bug fix: `peri_glue()` no longer modifies the calling environment

In `table.glue`, calling `table_glue()` injected an internal helper function
`..f` into the calling environment as a side effect. This was invisible in
most cases but could corrupt the caller's scope, and would raise an error if
the calling environment happened to be locked (as can occur in certain
`dplyr` contexts).

`peri_glue()` now evaluates in a disposable child environment, so nothing
leaks:

```{r}
check_leak <- function() {
  x <- 1.23456
  before <- ls(all.names = TRUE)
  peri_glue("{x}")
  after  <- ls(all.names = TRUE)
  setdiff(after, before)  # should be empty
}

check_leak()
```

## Bug fix: `format_decimal()` validates mark width

`table.glue` silently accepted multi-character decimal marks, which worked in
older R versions but was deprecated and is becoming an error in recent R.
`periglue` catches this at the point of specification rather than at
formatting time:

```{r, error = TRUE}
round_spec() |> format_decimal(mark = "decimal!")
```
